---
title: "ASU longitudinal data recoding"
output: html_document
date: '2022-09-18'
---

The wave 1 and wave 3 joined dataset - longitudinal section outputs

```{r}
library(tidyverse)
library(magrittr)
library(dplyr)
```

```{r}
wave1 <- read.csv("./../dataset_main/covid_pooled_public_w1b_1.1.0b7.csv", fileEncoding = 'UTF-8-BOM') #wave 1b 
wave2 <- read.csv("./../dataset_main/covidfuture_wave_2_public_Main_2.0.0b6.csv", fileEncoding = 'UTF-8-BOM') #wave 2
wave3 <- read.csv("./../dataset_main/covidfuture_wave_3_public_Main_3.0.0b6.csv") #wave 3
```

The choice selection for wave 1 and wave 3 have difference 

The ordinal answers of "w3_tr_freq_exp" for wave 3 are : "Never", "Once or twice/year", "A few times/year", "A few times/month", "Once/week or more"

The ordinal answers of "tr_freq_exp" for wave 1 are : 'Much less than before', 'Somewhat less than before', 'About the same', 'Somewhat more than before', 'Much more than before'

```{r}
df_wave1 <- list(wave1, wave3)     
df_wave1 %<>% reduce(inner_join, by='resp_id')

# replace the not responding value with NA 
df_wave1[df_wave1 == "Question not displayed to respondent"] <- NA
df_wave1[df_wave1 == "Variable not available in datasource"] <- NA
df_wave1[df_wave1 == "Seen but unanswered"] <- NA
df_wave1[df_wave1 == "Not selected"] <- NA
df_wave1[df_wave1 == "Nonmetropolitan"] <- NA

# remove the columns of missing data descriptions 
# remove the variable for showing attention ("trap_")
df_wave1 %<>%
  select(-contains("_miss"), -contains("trap_"), -contains("_trap"))


# ordinal encoding function 
encode_ordinal <- function(x) {
  #order <- unique(x)
  x <- as.numeric(factor(x, exclude = NA))
  x
}

# ordinal variables
Frequency    = c("Never", "A few times/year", "A few times/month", "Once/week", "A few times/week", "Every day")
Frequency1   = c("Never", "A few times/year", "A few times/month", "A few times/week", "Every day")
Frequency2   = c("Never", "Once or twice/year", "A few times/year", "A few times/month", "Once/week or more")
Frequency3   = c("Never", "A few times a year", "A few times a month", "A few times a week", "Every day")
Agreement    = c("Strongly disagree", "Somewhat disagree", "Neutral", "Somewhat agree", "Strongly agree")
Ld_change    = c("Much less", "Somewhat less", "About the same", "Somewhat more", "Much more")
Learn_change = c("Much worse", "Somewhat worse", "In some ways it is better and in some ways it is worse", "About the same", "Somewhat better", "Much better")
Prod_change  = c("Decreased significantly", "Decreased somewhat", "In some ways it has increased and in other ways it has decreased", "About the same", "Increased somewhat",
                 "Increased significantly")
Workload     = c("Yes, it has substantially decreased", "Yes, it has somewhat decreased", "No, it has stayed about the same", "Yes, it has somewhat increased", 
                 "Yes, it has substantially increased")
Compare      = c('Much less than before', 'Somewhat less than before', 'About the same', 'Somewhat more than before', 'Much more than before')
Shopping     = c('0', '1', '2 or 3', '4 or more')
Income       = c("Less than $10,000", "$10,000 to $14,999", "$15,000 to $24,999", "$25,000 to $34,999", "$35,000 to $49,999", "$50,000 to $74,999", "$75,000 to $99,999", 
                "$100,000 to $124,999", "$125,000 to $149,999", "$150,000 to $199,999", "$200,000 or more")
Incomechange = c('It has gone down a lot','It has gone down somewhat','It has stayed about the same', 'It has gone up somewhat', 'It has gone up a lot')
Education    = c("Some grade/high school", "Completed high school or GED", "Some college or technical school","Bachelor's degree(s) or some graduate school","Completed graduate degree(s)")
Risk         = c("Extremely low risk", "Low risk", "Medium risk", "High risk", "Extremely high risk")


# categorical variables
Mode = c(unique(df_wave1$pre_work_pri_mode))
Unit = c(unique(df_wave1$housingunit))
Mortgage = c(unique(df_wave1$tenure))




## Working section 
working <- df_wave1 %>%
    select(
         # id
         resp_id,       
         # employment  
         worker_pre,
         worker_now,
         jobchangejs,
         studentjs,    
         # work from home 
         contains("wfh"),
         workload,
         contains("bzvm"),
         contains("prod_"),       
         # study from home 
         ocla_pre,
         contains("class_sche"),
         learn_change,
         contains("learn_")

         )  %>%

    mutate(wfh_pre_freq  = factor(wfh_pre_freq, order=TRUE, levels = Frequency),
           wfh_freq_exp  = factor(wfh_freq_exp, order=TRUE, levels = Frequency),
           att_wfh_socint  = factor(att_wfh_socint, order=TRUE, levels = Agreement),
           att_wfh_hardmot  = factor(att_wfh_hardmot, order=TRUE, levels = Agreement),
           att_wfh_techfrus  = factor(att_wfh_techfrus, order=TRUE, levels = Agreement),
           workload = factor(workload, order=TRUE, levels = Workload),
           bzvm_pre = factor(bzvm_pre, order=TRUE, levels = Frequency1),
           bzvm_exp = factor(bzvm_exp, order=TRUE, levels = Frequency1),
           prod_change = factor(prod_change, order=TRUE, levels = Prod_change),
           learn_change = factor(learn_change, order=TRUE, levels = Learn_change)
          )

for(i in 2:ncol(working)) {# for-loop over columns
  working[ , i] <- encode_ordinal(working[ , i])
}






## Attitudes section - chaneged from wave 1
attitudes <-df_wave1 %>% 
  select(
         # id
         resp_id,        
         #all attitudes 
         contains("att_"),
         contains("risk_percp"),
         contains("enjoy_change")
          )%>%
  mutate(
    across(.cols = c(contains("att_")),
    .fns = ~factor(.x, order=TRUE, levels = Agreement)),
    across(.cols = c(contains("risk_percp")),
    .fns = ~factor(.x, order=TRUE, levels = Risk)),
    enjoy_change_YN = factor(enjoy_change_YN, order=TRUE, levels = c('No', 'Maybe', 'Yes'))) %>%

           
  mutate(across(.cols = -resp_id,
    .fns = encode_ordinal))  




## Commuting information 
commute_travel <-df_wave1 %>% 
  select(   
         # id
         resp_id, 
         
         # travel mode before, during and after covid 
         pre_work_pri_mode, 
         pre_sch_pri_mode, 
         now_work_pri_mode,
         apr_sch_pri_mode,
         scom_mode_exp, 
         wcom_mode_exp, 
         # days of commuting 
         pre_work_com_days,
         pre_sch_com_days,
         now_work_com_days,
         apr_sch_com_days,
         # travel time for commuting
         pre_work_pri_time,
         pre_sch_pri_time,
         now_work_pri_time,
         apr_sch_pri_time,
         # travel distance for commuting 
         pre_work_com_dist,
         pre_sch_com_dist,
         now_com_dist
         )%>%
  
  mutate(across(
    .cols = c(contains("mode")),
    .fns = ~factor(.x, order=TRUE, levels = Mode))) %>%
  mutate(across(.cols = c(contains("mode")),
    .fns = encode_ordinal)) 




## Transport change - the change from wave 1
transport_change <-df_wave1 %>%
  select(
         # id
         resp_id,
         bike,
         w3_bike,
         driver,
         # accessibility  
         # transit, active mobility, ride hail,telework change 
         contains("tr_freq")
         ) %>%
  
  mutate(across(
    .cols = c(contains("tr_freq_pre")),
    .fns = ~factor(.x, order=TRUE, levels = Frequency1))) %>%
  mutate(across(
    .cols = c(contains("w3_tr_freq_exp")),
    .fns = ~factor(.x, order=TRUE, levels = Frequency3))) %>%
  mutate(across(
    .cols = c(contains("tr_freq_exp_bikepers")),
    .fns = ~factor(.x, order=TRUE, levels = Compare))) %>%
  
  mutate(across(.cols = c(contains("tr_freq_pre"),contains("tr_freq_exp"),w3_bike,bike,driver),
    .fns = encode_ordinal))



## Shopping and Dinning
shopping_dinning <-df_wave1 %>%
  select(
         # id
         resp_id, 
         # shopping dining change 
         contains("shdi")
         ) %>%
  
  mutate(across(
    .cols = c(contains("shdi_pre"),contains("tr_freq_exp")),
    .fns = ~factor(.x, order=TRUE, levels = Frequency1))) %>%  
  
  mutate(across(
    .cols = c(contains("shdi_now")),
    .fns = ~factor(.x, order=TRUE, levels = Shopping))) %>% 
  
  mutate(across(.cols = c(contains("shdi")),
    .fns = encode_ordinal)) 




## demographic
demographic <- df_wave1 %>%
  select(
         # id
         resp_id,  
         # demographic 
         age, gender, hispanic, 
         contains("race_"),
         contains("disability_"),
         educ, # education 
         # geographic 
         ZIP_pre,
         ZIP_now,
         cbsa_code_pre,
         cbsa_code_now, 
         # housing information 
         housingunit, bedrooms, patyrd, hhsize, 
         contains("hhadults_"), # demographic information of adults number
         hhchild, 
         hhincome,
         income_impact,
         income_change,
         home_move, 
         home_move_exp,
         # vehicle 
         hhveh,
         contains("veh_"),
         # Internet
         tenure,
         contains("internet"),
         contains("devices"),
         contains("connect_change")
         ) %>%
    mutate(across(
    .cols = c(educ),
    .fns = ~factor(.x, order=TRUE, levels = Education))) %>% 
  
    mutate(educ = factor(educ, order=TRUE, levels = Education),
           housingunit = factor(housingunit, order=TRUE, levels = Unit),
           hhincome = factor(hhincome, order=TRUE, levels = Income),     
           income_change = factor(income_change, order=TRUE, levels = Incomechange),
           hhveh = factor(hhveh, order=TRUE, levels = c('0', '1', '2', '3', '4 or more')),
           tenure = factor(tenure, order=TRUE, levels = Mortgage)) %>% 
    mutate(across(
    .cols = c(contains("internet")),
    .fns = ~factor(.x, order=TRUE, levels = c("No", "I'm not sure", "Yes")))) %>% 
  
   mutate(across(.cols = c(contains("ZIP")|contains("cbsa"),hhveh),
     .fns = as.numeric))%>%

  mutate(across(.cols = c(where(is.character)|where(is.factor)),
    .fns = encode_ordinal)) 


## refine
demographic %<>%
  mutate(hispanic = hispanic-1,
       patyrd = patyrd-1,
       home_move = home_move-1,
       veh_sold = veh_sold -1,
       veh_purchase = veh_purchase-1,
       internet_con = internet_con-1,
       mob_internet = mob_internet-1)%>%
  mutate(home_move = replace(home_move, home_move == 2, 1))

transport_change %<>%
  mutate(w3_bike = w3_bike-1,
         bike = bike-1,
         driver = driver-1)

working %<>%
  mutate(worker_pre = worker_pre-1,
         worker_now = worker_now-1,
         jobchangejs = jobchangejs-1,
         studentjs = studentjs-1,
         wfh_pre = wfh_pre-1,
         wfh_now = wfh_now-1,
         wfh_expect = wfh_expect-1)

attitudes %<>%
  mutate(enjoy_change_YN = enjoy_change_YN-1)#%>%
  #mutate(enjoy_change_YN = replace(enjoy_change_YN, enjoy_change_YN == 2, 1))

#put all data frames into list
df <- list(demographic, working, commute_travel, transport_change,shopping_dinning, attitudes)      
#merge all data frames together
df %<>% reduce(full_join, by='resp_id')

#write.csv(df, "datasecondwave.csv",row.names = FALSE, na = ".")


```


Getting the selected questions and variables of cycling for wave 1 and wave 3 samples. 



```{r}

data_bike <- df %>%
 transmute(
    tr_freq_pre_bikepers,
    tr_freq_now_bikepers,
    tr_freq_exp_bikepers,
    w3_tr_freq_now_8,
    w3_tr_freq_exp_8,
    educ,
    age,
    hhincome,
    hhveh,
    bike,
    #w3_bike,
    bzvm_pre,
    race_1,
    disability_7,
    
    # covid attitudes from wave 1 
    att_covid_selfsevere,
    att_covid_friendssevere,
    att_covid_stayhome,
    att_covid_stayhome_norm,
    risk_percp_1,
    risk_percp_2,
    risk_percp_3,
    risk_percp_4,
    risk_percp_5,
    risk_percp_6,
    att_covid_overreact,
    att_covid_economic,
    
    # Life convenience attitudes from wave 1 
    att_shop_online,
    att_wfh_likewfh.y,
    att_tech_videobusiness,
    att_tech_videofriends,
    att_tech_hscolonline,
    
    # Environmental awareness attitudes from wave 1 
    att_environ_lifst,
    att_environ_lesspolluting,
    att_environ_plane,
    
    # Pre covid attitudes to being social attitudes from wave 1 
    att_socit_safetransit,
    att_socit_outside,
    att_socit_others,
    
    # outsides Index attitudes from wave 1 
    att_shop_chore,
    att_wfh_socint.y,
    att_wfh_hardmot.y,
    att_lifst_transition,
  
    freq_change = w3_tr_freq_now_8 - tr_freq_now_bikepers,
    
    # covid attitudes from wave 3
    w3_att_covid_1,
    w3_att_covid_2,
    w3_att_covid_3,
    w3_att_covid_6,
    w3_risk_percp_1,
    w3_risk_percp_2,
    w3_risk_percp_3,
    w3_risk_percp_4,
    w3_risk_percp_5,
    w3_risk_percp_6,
    w3_att_covid_4,
    w3_att_covid_5,
    
    # Life convenience attitudes from wave 3 
    w3_att_shop_3,
    w3_att_wfh_correct_shdi_5,
    w3_att_tech_1,
    w3_att_tech_2,
    w3_att_tech_3,
    
    # Environmental awareness attitudes from wave 3
    w3_att_environ_1,
    w3_att_environ_2,
    w3_att_environ_3,
    
    # Pre covid attitudes to being social
    #w3_att_socit_safetransit,
    #w3_att_socit_outside,
    #w3_att_socit_others,
    
    # outsides Index attitudes from wave 3
    w3_att_shop_5,
    w3_att_wfh_correct_shdi_3,
    w3_att_wfh_correct_shdi_4,
    w3_att_lifst_3
    )

data_bike$freq_change[data_bike$freq_change > 3] <- 5
data_bike$freq_change[data_bike$freq_change > 0 & data_bike$freq_change <= 3] <- 4
data_bike$freq_change[data_bike$freq_change == 0] <- 3
data_bike$freq_change[data_bike$freq_change >= -3 & data_bike$freq_change < 0] <- 2 
data_bike$freq_change[data_bike$freq_change < -3] <- 1



write.csv(data_bike, "./data_recoding_output/data_bike_wave13.csv",row.names = FALSE, na = ".")
```


Getting the selected questions and variables of walking for wave 1 and wave 3 samples. 

```{r}
data_walk <- df %>%
  transmute(
    tr_freq_pre_walk,
    tr_freq_now_walk,
    tr_freq_exp_walk,
    w3_tr_freq_now_6,
    w3_tr_freq_exp_6,
    educ,
    age,
    hhincome,
    hhveh,
    bike,
    #w3_bike,
    bzvm_pre,
    race_1,
    disability_7,
    
    # covid attitudes from wave 1 
    att_covid_selfsevere,
    att_covid_friendssevere,
    att_covid_stayhome,
    att_covid_stayhome_norm,
    risk_percp_1,
    risk_percp_2,
    risk_percp_3,
    risk_percp_4,
    risk_percp_5,
    risk_percp_6,
    att_covid_overreact,
    att_covid_economic,
    
    # Life convenience attitudes from wave 1 
    att_shop_online,
    att_wfh_likewfh.y,
    att_tech_videobusiness,
    att_tech_videofriends,
    att_tech_hscolonline,
    
    # Environmental awareness attitudes from wave 1 
    att_environ_lifst,
    att_environ_lesspolluting,
    att_environ_plane,
    
    # Pre covid attitudes to being social attitudes from wave 1 
    att_socit_safetransit,
    att_socit_outside,
    att_socit_others,
    
    # outsides Index attitudes from wave 1 
    att_shop_chore,
    att_wfh_socint.y,
    att_wfh_hardmot.y,
    att_lifst_transition,
  
    freq_change = w3_tr_freq_now_6 - tr_freq_now_walk,
    

    # covid attitudes from wave 3
    w3_att_covid_1,
    w3_att_covid_2,
    w3_att_covid_3,
    w3_att_covid_6,
    w3_risk_percp_1,
    w3_risk_percp_2,
    w3_risk_percp_3,
    w3_risk_percp_4,
    w3_risk_percp_5,
    w3_risk_percp_6,
    w3_att_covid_4,
    w3_att_covid_5,
    
    # Life convenience attitudes from wave 3 
    w3_att_shop_3,
    w3_att_wfh_correct_shdi_5,
    w3_att_tech_1,
    w3_att_tech_2,
    w3_att_tech_3,
    
    # Environmental awareness attitudes from wave 3
    w3_att_environ_1,
    w3_att_environ_2,
    w3_att_environ_3,
    
    # Pre covid attitudes to being social
    #w3_att_socit_safetransit,
    #w3_att_socit_outside,
    #w3_att_socit_others,
    
    # outsides Index attitudes from wave 3
    w3_att_shop_5,
    w3_att_wfh_correct_shdi_3,
    w3_att_wfh_correct_shdi_4,
    w3_att_lifst_3
    )
data_walk$freq_change[data_walk$freq_change > 3] <- 5
data_walk$freq_change[data_walk$freq_change > 0 & data_walk$freq_change <= 3] <- 4
data_walk$freq_change[data_walk$freq_change == 0] <- 3
data_walk$freq_change[data_walk$freq_change >= -3 & data_walk$freq_change < 0] <- 2 
data_walk$freq_change[data_walk$freq_change < -3] <- 1


write.csv(data_walk, "./data_recoding_output/data_walk_wave13.csv",row.names = FALSE, na = ".")
```

